Vagueness \& Rationality

### 2018-08-18 springer/current/e1.tex
added a full table of instructiosn for D-Exp-2

@@

Examples of crisp and vague versions of the numerical and verbal instructions follow: the examples assume the array (6,15,24) and reference to the smaller number of dots, such that 6 was classified as the expected response; 15 as the borderline response; and 24 as the incorrect response. In the {\em vague numerical} condition we used \emph{Choose a square with about 10 dots}. None of the squares contained 10 dots. 10 is slightly closer to 6 than to 15, justifying 6 as the best response and 15 as the borderline response. In the {\em vague verbal} condition we used \emph{Choose a square with few dots}. In the {\em crisp numerical} condition we used \emph{Choose the square with 6 dots}, and one square always did contain the number mentioned. For {\em crisp verbal}, we used \emph{Choose the square with the fewest dots}.

>>>>

\marginnote{C\_Exp\_1}

@@

Examples of crisp and vague versions of the numerical and verbal instructions are given next: the examples assume the array (6,15,24) and reference to the smaller number of dots, such that 6 was classified as the expected response; 15 as the borderline response; and 24 as the incorrect response. In the {\em vague numerical} condition we used \emph{Choose a square with about 10 dots}. None of the squares contained 10 dots. 10 is slightly closer to 6 than to 15, justifying 6 as the best response and 15 as the borderline response. In the {\em vague verbal} condition we used \emph{Choose a square with few dots}. In the {\em crisp numerical} condition we used \emph{Choose the square with 6 dots}, and one square always did contain the number mentioned. For {\em crisp verbal}, we used \emph{Choose the square with the fewest dots}.

### 2018-08-18 springer/current/e2.tex
added a placeholder citation for Levy article on main effects and interactions
added a full table of instructiosn for D-Exp-2

The main aim of Experiment 3 was to see whether vagueness would exert beneficial effects when all conditions used numerals in the instructions, and when there were vague and crisp versions of the instructions for both comparison and matching strategies. The main changes from Experiment 2 were that the human selection task was explicitly controlled (i.e., in whether it amounted to matching or comparison), and that all conditions were constrained to mention a number. We used the same arrays as in Experiment 2. Table \ref{Instructions for e3} shows the instructions for each condition. Note the difference between ``fewer than 20'' and ``far fewer than 20'': whereas the former cannot have borderline cases (i.e., for each number it is clear whether the number is smaller than 20 or not), the latter can.

\begin{table}
\centering
\caption{Experiment 3: Instructions arranged by condition for stimuli with $(6,15,24)$ dots and instructions indicating the smaller numbers} 
\label{Instructions for e3}
\begin{tabular}{cccc}
\hline\noalign{\smallskip}
Instruction	format          &Vagueness					&	Selection task	& Instruction									\\
\noalign{\smallskip}\hline\noalign{\smallskip}
\multirow{4}{*}{numeric}	&\multirow{2}{*}{crisp} 	&	matching		& Choose a square with 6 dots 					\\ 
							&							&	comparison 		& Choose a square with fewer than 20 dots 		\\
                             \cline{2-4}
							&\multirow{2}{*}{vague} 	&	matching 		& Choose a square with about 10 dots 			\\ 
							&							&	comparison 		& Choose a square with far fewer than 20 dots 	\\ 
\noalign{\smallskip}\hline
\end{tabular}
\end{table}%

\subsection{Hypotheses} % E 3
\noindent For Experiment 3, we hypothesised:

@@

\subsection{Experimental Method} % E 3

38 participants were recruited. We used a 2 x 2 factorial manipulation of vagueness and selection task (see Table \ref{Instructions for e3}).
On each trial an instruction was presented: participants pressed a key to dismiss the instruction, when the dot arrays were presented until the participant responded, and the response time and choice were recorded.

\subsection{Results} % E 3

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{figures/e3-rtplot-1.pdf}
\caption{Mean response times by condition for Experiment 3 where all instructions were numeric}
\label{resultse3}
\end{figure}

>>>>

\marginnote{D\_Exp\_2}

The main aim of Experiment 2 was to see whether vagueness would exert beneficial effects when all conditions used numerals in the instructions, and when there were vague and crisp versions of the instructions for both comparison and matching strategies. The main changes from Experiment 2 were that the human selection task was explicitly controlled (i.e., in whether it amounted to matching or comparison), and that all conditions were constrained to mention a number. We used the same arrays as in Experiment 2. Table \ref{Instructions for e2} shows the instructions for each condition. Note the difference between ``fewer than 20'' and ``far fewer than 20'': whereas the former cannot have borderline cases (i.e., for each number it is clear whether the number is smaller than 20 or not), the latter can.

\begin{table}
\centering
\caption{Experiment 2: Instructions arranged by condition. The instructions given in the table started with ``Choose a square with \ldots"} 
\label{Instructions for e2}
\begin{tabular}{cccll}
\hline\noalign{\smallskip}
Item & Quantity & Selection & Crisp & Vague \\ 
\noalign{\smallskip}\hline\noalign{\smallskip}
06:15:24 & Small & Comparison & fewer than 20 dots & far fewer than 20 dots \\ 
06:15:24 & Small & Matching & 6 dots & about 10 dots \\ 
06:15:24 & Large & Comparison & more than 10 dots & far more than 10 dots \\ 
06:15:24 & Large & Matching & 24 dots & about 20 dots \\ 
\noalign{\smallskip}\hline\noalign{\smallskip}
16:25:34 & Small & Comparison & fewer than 30 dots & far fewer than 30 dots \\ 
16:25:34 & Small & Matching & 16 dots & about 20 dots \\ 
16:25:34 & Large & Comparison & more than 20 dots & far more than 20 dots \\ 
16:25:34 & Large & Matching & 34 dots & about 30 dots \\ 
\noalign{\smallskip}\hline\noalign{\smallskip}
26:35:44 & Small & Comparison & fewer than 40 dots & far fewer than 40 dots \\ 
26:35:44 & Small & Matching & 26 dots & about 30 dots \\ 
26:35:44 & Large & Comparison & more than 30 dots & far more than 30 dots \\ 
26:35:44 & Large & Matching & 44 dots & about 40 dots \\ 
\noalign{\smallskip}\hline\noalign{\smallskip}
36:45:54 & Small & Comparison & fewer than 50 dots & far fewer than 50 dots \\ 
36:45:54 & Small & Matching & 36 dots & about 40 dots \\ 
36:45:54 & Large & Comparison & more than 40 dots & far more than 40 dots \\ 
36:45:54 & Large & Matching & 54 dots & about 50 dots \\ 
\noalign{\smallskip}\hline
\end{tabular}
\end{table}

\subsection{Hypotheses} % D_Exp_2
\noindent For Experiment 2, we hypothesised:

@@

\subsection{Experimental Method} % D_Exp_2

38 participants were recruited. We used a 2 x 2 factorial manipulation of vagueness and selection task (see Table \ref{Instructions for e2}).
On each trial an instruction was presented: participants pressed a key to dismiss the instruction, when the dot arrays were presented until the participant responded, and the response time and choice were recorded.

\subsection{Results} % D_Exp_2

We used the method proposed by \citet{Levy:MainEffectsInteractions} to test for the main effect of Vagueness in the presence of its higher-order interactions.

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{figures/e3-rtplot-1.pdf}
\caption{Mean response times by condition for Experiment 2 where all instructions were numeric}
\label{resultse3}
\end{figure}
